---
title: "Lecture 1"
author: "Han Zhang"
date: "2/1/2021"
output: pdf_document
---
# create population
For learning purposes, we generate a pseudo population. 
The population is 100,000 units following Beta distribution.

Note that the population distribution is not Normal.

Note that in real life, you don't observe the population. 
Here we assume we know them.
It's a common way to learn about statistics and design methods.

$E(X) = mean(population)$
```{r}
population <- rbeta (100000, 1, 2)
plot (density (population))
# population mean
mean(population) 
# population variance
var(population)

```

# create sample

We create *one* survey of 1000 sampled observations from the population.
$X_1,...,X_{1000}$

We calculated sample mean $\bar{X}$ and $\hat{V}(X)$
Compare these estimates with population quantities. What do you find?

```{r}
survey = sample(population, size = 1000)
plot (density (survey))
# sample mean as estimator of population mean
est.mean = mean(survey)

# unbiased estimator of population variance
est.var = 1000/(1000-1) * var(survey)

```

# Normally-approximated confidence intervals

First we need to estimate standard error of the sample mean

We know it is

$\hat{V}(\bar{X}) = \frac{\hat{V}(X)}{n}$. We have calculated $\hat{V}(X)$ in the previous code chunk.

```{r}
sample_mean = mean(survey)

#standard error of the sample mean
ss = sqrt(est.var/1000)

print ("point estimate of mean")
print (sample_mean)
print ("95% normal-approximated confidence interval of mean")
c(sample_mean - 1.96 * ss, sample_mean + 1.96 * ss)

```


# Law of Large Numbers
We vary the size of survey, from 100 to 100000, with an interval of 1000.

In other words, $n$ changes from 100, 500, 1100, 1600, 2100, and so on.

As the plot shows, when $n$ increases, sample mean approaches population mean.

```{r}
survey_mean_list  <- c()
survey_times <- seq(100, 100000, 500)
for (n in survey_times ){
  survey <- sample(population, size = n, replace = F)
  survey_mean_list  <- c(survey_mean_list , mean(survey))
}

plot(survey_times, survey_mean_list , xlab = "number of samples", ylab = "sample mean")
abline(h = mean(population), col = "red")
```

#  Central limit theorem

This time, we let survey size to be fixed at 1000.

We explore three cases:
- case 1: we create 10 surveys, each of size 1000, from the population. Plot color is red.
- case 2: we create 10000 surveys, each of size 1000, from the population. Plot color is blue.



```{r}

# Now we show the central limit theorem
survey_times <- c(10,  10000)
col <- c("red",  "blue") #"green",
i = 1
for (n in survey_times ){
  print (n)
  # central limit theorem talks about distribution of the sample mean
  # we cannot calculate the distribution for a single survey, so we draw sample multiple times
  survey_mean_list <- c()
  for (m in 1:n)
  {
    survey <- sample(population, size = 1000, replace = TRUE)
    survey_mean_list <- c(survey_mean_list, mean(survey))
  }
  # sample_mean_list_standard <- survey_mean_list  - mean(population)
  plot(density (survey_mean_list),  col = col[i], xlim = c(0.3, 0.38), ylim  = c(0,60), xlab = "")
  abline(v = mean(population), col = "black")
  par(new = T)
  i = i  + 1
}


```


## bootstrap confidence interval for sample mean

Bootstrap is slower, but needs no Central Limit Theorem.

NOTE: sample *with* replacement is the key

```{r}
print (length(sample))
bootstrap_means = c() # store bootstrapped means
for (i in 1:1000){
  # resample from the sample with replacement
  boot_data <- sample(survey, length(survey), replace = T)
  boot_mean <- mean(boot_data)
  bootstrap_means <- c(bootstrap_means, boot_mean)
}

# then simple quantile function to 
print ("point estimate of mean")
print (sample_mean)

print ("point estimate of mean")
print (mean(bootstrap_means))

print ("normal confidence interval")
print (c(sample_mean - 1.96 * ss, sample_mean + 1.96 * ss))

print ("95% bootstrap confidence interval of mean")
conf = quantile(bootstrap_means, c(0.025, 0.975))
print (conf)
plot(density(bootstrap_means), main = "confidence interval")
abline(v = conf[1], col = "red")
abline(v = conf[2], col = "red")
abline(v = sample_mean - 1.96 * ss, col = "green")
abline(v = sample_mean + 1.96 * ss, col = "green")


```
