---
title: "Lec5"
author: "Han Zhang"
date: "2024-03-17"
output: html_document
---

following
https://grf-labs.github.io/grf/

```{r}
library(grf)

# Generate data.
n <- 2000
p <- 10
X <- matrix(rnorm(n * p), n, p)
X.test <- matrix(0, 101, p)
X.test[, 1] <- seq(-2, 2, length.out = 101)
X.test[, 2] <- rnorm(101)

# Train a causal forest.
## W is your treament variable!!
W <- rbinom(n, 1, 0.4 + 0.2 * (X[, 1] > 0))
Y <- pmax(X[, 1], 0) * W + X[, 2] + pmin(X[, 3], 0) + rnorm(n)


data <- data.frame(X,W,Y)
```

How would you do without GRF
```{r}
m1 <- lm(Y ~ W * X1, data)
m2 <- lm(Y ~ W * X2, data)
## add more interactions?

m2 <- lm(Y ~ W * X1 + W*X2, data)
## keep going??
```


## by default, honest prediction is used


```{r}
tau.forest <- causal_forest(X, Y, W)

# Estimate treatment effects for the training data using out-of-bag prediction.
tau.hat.oob <- predict(tau.forest)
hist(tau.hat.oob$predictions)
```

## this used train/test split 

So you know very clearly what you are doing (that which test is *not* used during the training process)


X.test never entered into fitting /trainig process
But we used it to calculate the effect
```{r}
# Estimate treatment effects for the test sample.
tau.hat <- predict(tau.forest, X.test)
## plot X1 with treatment effects
plot(X.test[, 1], tau.hat$predictions, ylim = range(tau.hat$predictions, 0, 2), xlab = "x1", ylab = "tau", type = "l")
lines(X.test[, 1], pmax(0, X.test[, 1]), col = 2, lty = 2)
## plot X2 with treatment effects
plot(X.test[, 2], tau.hat$predictions, ylim = range(tau.hat$predictions, 0, 2), xlab = "x2", ylab = "tau", type = "l")
lines(X.test[, 2], pmax(0, X.test[, 2]), col = 2, lty = 2)

```



# Estimate the conditional average treatment effect on the full sample (CATE).
```{r}
average_treatment_effect(tau.forest, target.sample = "all")

# Estimate the conditional average treatment effect on the treated sample (CATT).
average_treatment_effect(tau.forest, target.sample = "treated")

# Add confidence intervals for heterogeneous treatment effects; growing more trees is now recommended.
tau.forest <- causal_forest(X, Y, W, num.trees = 4000)
tau.hat <- predict(tau.forest, X.test, estimate.variance = TRUE)
sigma.hat <- sqrt(tau.hat$variance.estimates)


plot(X.test[, 1], tau.hat$predictions, ylim = range(tau.hat$predictions + 1.96 * sigma.hat, tau.hat$predictions - 1.96 * sigma.hat, 0, 2), xlab = "x", ylab = "tau", type = "l")
lines(X.test[, 1], tau.hat$predictions + 1.96 * sigma.hat, col = 1, lty = 2)
lines(X.test[, 1], tau.hat$predictions - 1.96 * sigma.hat, col = 1, lty = 2)
lines(X.test[, 1], pmax(0, X.test[, 1]), col = 2, lty = 1)
```


